---
title: "Reproducible Research - Project 1"
author: "Angus Mackenzie"
date: "30/05/2021"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=TRUE}
## Setup
options(scipen=999)
knitr::opts_chunk$set(echo = TRUE)

## Libraries
library(tidyverse)
library(lubridate)
library(ggrepel)
library(RColorBrewer)

```

## Loading and preprocessing the data

Here is the data, downloaded and unzipped from the course website

```{r}

## Data
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "steps.zip")
unzip(zipfile = "steps.zip")

dfstep <- read.csv("activity.csv") %>%
    mutate(date = ymd(date))

```

## What is mean total number of steps taken per day?

```{r,dpi=500,fig.height=3,fig.width=7,fig.align='center'}

## Steps per day
dfstep_pd <- dfstep %>%
  group_by(date) %>%
  summarise(daily_steps = sum(steps, na.rm = TRUE))

daily_steps_mean <- mean(dfstep_pd$daily_steps, na.rm = TRUE)
daily_steps_median <- median(dfstep_pd$daily_steps, na.rm = TRUE)

## Number of intervals
n_interval <- length(unique(dfstep$interval))

## Graph
dfstep_pd %>%
  ggplot(aes(x = daily_steps)) +
  geom_histogram(fill = 'lightblue', binwidth = n_interval) +
  theme_bw()


```

* The mean number of steps per day is `r daily_steps_mean`
* The median number of steps per day is `r daily_steps_median`

## What is the average daily activity pattern?
```{r,dpi=500,fig.height=3,fig.width=7,fig.align='center'}

## Time series
dfstep_ts <- dfstep %>%
  group_by(interval) %>%
  summarise(interval_steps = mean(steps, na.rm = TRUE))

interval_steps_max <- dfstep_ts$interval[which.max(dfstep_ts$interval_steps)]

## Graph
dfstep_ts %>% 
  ggplot(aes(x = interval, y = interval_steps)) +
  geom_line(color = 'blue') +
  theme_bw()


```

The 5 minute interval with the greatest average steps is `r interval_steps_max`

## Imputing missing values
```{r,dpi=500,fig.height=3,fig.width=7,fig.align='center'}

## Missing values
dfstep_na <- dfstep %>%
  filter(is.na(steps))

missing_step_count <- nrow(dfstep_na)

## Fill with mean
dfstep_fill <- dfstep %>%
  mutate(fill_steps = ifelse(is.na(steps), dfstep_ts$interval_steps[match(dfstep_ts$interval, interval)], steps)) %>%
  group_by(date) %>%
  summarise(fill_steps = sum(fill_steps, na.rm = TRUE))

fill_steps_mean <- mean(dfstep_fill$fill_steps, na.rm = TRUE)
fill_steps_median <- median(dfstep_fill$fill_steps, na.rm = TRUE)
fill_total_diff <- sum(dfstep_fill$fill_steps) - sum(dfstep_pd$daily_steps)

## Graph
dfstep_fill %>%
  ggplot(aes(x = fill_steps)) +
  geom_histogram(fill = 'lightblue', binwidth = n_interval) +
  theme_bw()
```

* Filling in missing values with the interval mean gives a different mean number of steps per day of `r fill_steps_mean`
* Filling in missing values with the interval mean gives a different median number of steps per day of `r fill_steps_median`
* The total steps taken after filling in missing values is `r fill_total_diff` more than before.

## Are there differences in activity patterns between weekdays and weekends?

```{r,dpi=500,fig.height=5,fig.width=7,fig.align='center'}

## Separate weekdays
vweekday <- c("Mon","Tue","Wed","Thu","Fri")

dfstep_week <- dfstep %>%
  mutate(iday = ifelse(weekdays(date, abbreviate = TRUE) %in% vweekday, "Weekday", "Weekend")) %>%
  group_by(interval, iday) %>%
  summarise(interval_steps = mean(steps, na.rm = TRUE))

## Graph
dfstep_week %>% 
  ggplot(aes(x = interval, y = interval_steps)) +
  facet_wrap(~iday, ncol = 1) +
  geom_line(color = 'blue') +
  theme_bw()
```

